---
title: "Eddie Module on Nutrient Loading"
output: html_notebook
---

Note this is essentially the same thing as the water quality module and I think there is an issue with the set up as it is not correct


# Install the basic libraries
```{r install libraries}
# install.packages("tidyverse")
# install.packages("readxl")
# install.packages("janitor")
# install.packages("lubridate")
# install.packages("broom")
# install.packages("patchwork")
# install.packages("skimr")
```

# Load libraries    
Load the libraries each time you run a script or program   
```{r load libraries, message=FALSE, warning=FALSE}
library(tidyverse) # loads a lot of libraries in this one to add more functionality
library(readxl) # allows you to read in excel files
library(janitor) # cleans up column names and removes empty columns if wanted
library(lubridate) # allows easy conversion of varaibles to date format
library(broom) # cleans up output for easy presentation
library(patchwork)
library(skimr)
```

           
# Read in files
Decar River in Palo Iowa 15 mintue data
```{r}
decar.df <- read_excel("../data/nutrient_loading_module_data.xlsx",
                    sheet = "decar_river_palo_iowa")
```


# Recalculate columns for reproducibility
The conversion of cubic feet per second to liters required CFS to be multiplied by 28.136
```{r}
decar.df <- decar.df %>% 
  mutate(time_int_load_kgn = discharge_cfs * nitrate_n_mgl * 28.136 * 15 * (60/1000000),
         time_int__q_l = discharge_cfs *28.136*15*60)
```


# Plot of nitrate versus time
```{r}
decar.df %>% 
  ggplot(aes(x=datetime, y =nitrate_n_mgl)) +
  geom_line() +
  geom_point()
```

```{r}
decar.df %>% 
  ggplot(aes(x=datetime, y =discharge_cfs)) +
  geom_line() +
  geom_point()
```

# Statistics of nitrate data
```{r}
skim(decar.df$ nitrate_n_mgl)
```

Add a new column to the right of the concentration data called “Rank” (R).  In this column, the highest concentration value has a rank equal to 1, the next highest concentration has a rank of 2, and so forth for the entire dataset.  There should be 13,897 data points in this dataset.  This means the total number (n) of this dataset is 13,897.       
    
Add a third column to the right called Probability (p).  This will be the likelihood that a concentration of a given Rank (R) will occur within the dataset.  Calculate the probability by dividing the rank by the total number (n) plus one.  Type this equation in a new column and fill it down for the entire dataset.  p=R/(n+1)      
      
Finally, calculate a fourth column showing the Percent Probability (%), by multiplying the probability number by 100.  This is the probability that a given concentration occurred during any 15 second period between April and September, 2014.       


# Create a probability dataset
This will create a probability column in a new dataset called pr
```{r}
prob.df <- decar.df %>% 
  filter(!is.na(nitrate_n_mgl)) %>% 
  arrange(desc(nitrate_n_mgl)) %>% 
  mutate(rank  = 1:n()) %>% 
  mutate(prob = rank / (n()+1)) %>%
  mutate(pct_prob = prob * 100)
```


```{r}
prob.df %>% 
  ggplot(aes(x=pct_prob, y = nitrate_n_mgl)) +
  geom_point() +
  geom_hline(yintercept = 10, color="red", size = 2, alpha = .5) +
  annotate(geom="text",
                x=25,
                y=10.5,
                label="EPA Drinking Water Limit",
                fontface="bold")
```




